---
title: "Assignment 1 Final Output"
output: 
    pdf_document:
        toc: true
---

This a summary of the descriptive statistics from Diego and Isabelle's first assignment for the Collaborative Social Science Data Analysis class.

\pagebreak

# Data Set 1: Average Yearly Temperatures in New Haven 1912-1971

## Introduction to the DataSet
The core R data set nhtemp, titled "Average Yearly Temperatures in New Haven" is a time series of 60 observations recording the mean annual temperature in degrees Fahrenheit in New Haven, Connecticut, between 1912 and 1971.

## Descriptive Statistics

If we look at the summary statistics, we learn that the average temperature has fallen within a small range for the duration of the 60 years. Otherwise, it tells us very little about the actual climate or any change in climate over time, as the temperatures are aggregated by year and not by month or season.

```{r}
summary(nhtemp)
```

However, if we look at the data over time, we can see a general upwards trend in the average temperatures, although they vary considerably from one year to the next.

```{r, echo=FALSE}
plot(nhtemp,
    main="Average Yearly Temperature 1912-1971",
    ylab="Temperature (Fahrenheit)")
```


\pagebreak

The histogram shows us that most years cluster towards an average temperature of 50-52.

```{r, echo=FALSE}
hist(nhtemp, 
     main="Frequency of Average Yearly Temperatures", 
     xlab="Temperature")
```

\pagebreak

# Data Set 2: Violent Crime Rates by US State

## Introduction to the DataSet

The dataset USArrests contains statistics, in arrests per 100,000 residents for assault, murder, and rape in each of the 50 US states in 1973. It also provides the percent of the population living in urban areas.

There are 50 observations on the 4 following variables:

-Murder (murder arrests per 100,000)

-Assault (assault arrests per 100,000)

-UrbanPop (percent urban population per state)

-Rape (rape arrests per 100,000)

##Descriptive Statistics

##Summary statistics

```{r}
summary(USArrests$Murder)
```

Firstly, we use the summary function to have a look at the values of the mean and the median (measures of central tendency) for each of the 4 variables. For murder arrests per 100,000 population, the mean is 7.788 and the median is 7.250.

```{r}
summary(USArrests$Assault)
```

For assault arrests per 100,000 population, the mean is 170.8 and the median is 159. 

```{r}
summary(USArrests$UrbanPop)
```

For percent of urban population in each state, the mean is 65.54% and the median is 66%. 

```{r}
summary(USArrests$Rape)
```

For rape arrests per 100,000 population, the mean is 21.23 and the median is 20.10. 

\pagebreak

##Histograms

Secondly, we generate histograms in order to have a general idea about the distribution of the 4 variables contained in our dataset. For murder arrests per 100,000 population, the histogram tells us that the largest group of US states (12 states) has between 2 and 5 arrests for murder per 100,000 residents. There is only 1 state that has less than 2 arrests for murder per 100,000 residents and more than 16 states have between 6 and 10 arrests per 100,000 residents (two bars of the histogram).
For assault arrests per 100,000 residents, the histogram tells us that the largest group of states (12 states) has between 100 and 150 arrests for assault per 100,000 residents. The second largest group of states (10 states) has between 250 and 300 arrests for assault per 100,000 residents. On the other hand, while 3 states only have between 0 and 50 arrests for assault per 100,000 residents, 2 states have between 300 and 350. 
For the share of urban population in each US state, the largest group of states (12 states) has between 60 and 70 % of their population living in urban areas. Likewise, 10 states have between 70 and 80 % of their population living in urban areas. On the other hand, 2 states have less than 40 % of their population living in urban areas, and 1 state has more than 90% of urban population. When it comes to rape arrests per 100,000 residents, the distribution appears to be right skewed since a very small number of US states have more than 30 arrests for rape per 100,000 residents. The largest number of states (12 states) have between 15 and 20 arrests for rape per 100,000 residents.  

```{r, echo=FALSE}
hist(USArrests$Murder,
     main = "Arrests for Murder across US states",
     ylab = "Number of US states",
     xlab = "Arrests for Murder per 100k residents")
```

\pagebreak

```{r, echo=FALSE}
hist(USArrests$Assault,
     main = "Arrests for Assault across US states",
     ylab = "Number of US states",
     xlab = "Arrests for Assault per 100k residents")
```

```{r, echo=FALSE}
hist(USArrests$UrbanPop,
     main = "Share of Urban Population across US states",
     ylab = "Number of US states",
     xlab = "Share of Urban Population (in %)")
```

\pagebreak

```{r, echo=FALSE}
hist(USArrests$Rape,
     main = "Arrests for Rape across US states",
     ylab = "Number of US states",
     xlab = "Arrests for Rape per 100k residents")
```

##Boxplots

Now we make use of boxplots in order to provide a graphical representation of where the mean, median and interquartile range of our three crime variables are. 

```{r, echo=FALSE}
boxplot(USArrests$Murder,
        main = "Murder Arrests",
        ylab = "Arrests per 100k Inhabitants")
```

```{r, echo=FALSE}
boxplot(USArrests$Assault,
        main = "Assault Arrests",
        ylab = "Arrests per 100k Inhabitants")
```

```{r, echo=FALSE}
boxplot(USArrests$Rape,
        main = "Rape Arrests",
        ylab = "Arrests per 100k Inhabitants")
```

\pagebreak

##Correlations

Finally, scatter plots help us assess whether there is correlation between our main variables of interest. 

```{r, echo=FALSE}
library(ggplot2)
ggplot2::ggplot(USArrests, aes(Murder, Assault)) + geom_point() + geom_smooth() + theme_bw() + ggtitle("Murder arrests-Assault arrests Correlation") +
  xlab("Murder arrests per 100k habitants") +
  ylab("Assault arrests per 100k habitants")
```

```{r}
cor.test(USArrests$Murder, USArrests$Assault)
```

As both the scatter plot and the Pearson correlation coefficient indicate, there seems to be a strong positive correlation between murder arrests and assault arrests. 

\pagebreak

```{r, echo=FALSE}
library(ggplot2)
ggplot2::ggplot(USArrests, aes(Murder, Rape)) + geom_point() + geom_smooth() + theme_bw() + ggtitle("Murder arrests-Rape arrests Correlation") +
  xlab("Murder arrests per 100k habitants") +
  ylab("Rape arrests per 100k habitants")
```

```{r}
cor.test(USArrests$Murder, USArrests$Rape)
```

There also seems to be correlation between murder arrests and rape arrests, although slightly weaker than in the previous case. 

\pagebreak

```{r, echo=FALSE}
library(ggplot2)
ggplot2::ggplot(USArrests, aes(Assault, Rape)) + geom_point() + geom_smooth() + theme_bw() + ggtitle("Assault arrests-Rape arrests Correlation") +
  xlab("Assault arrests per 100k habitants") +
  ylab("Rape arrests per 100k habitants")
```

```{r}
cor.test(USArrests$Assault, USArrests$Rape)
```

The correlation between assault arrests and rape arrests is also positive.   
In conclusion, there seems to be a strong positive correlation between the three variables of interest. Nevertheless, the strongest correlation is the one between murder arrests and assault arrests. 