---
title: "Assignment 1 Final Output"
output: pdf_document
---

This a summary of the descriptive statistics from Diego and Isabelle's first assignment for the Collaborative Social Science Data Analysis class.

## *Table of Contents*
#[Data Set 1: Average Yearly Temperatures in New Haven 1912-1971](#dataset1)
#[Data Set 2: Violent Crime Rates by US State](#dataset2)

\pagebreak

#Data Set 1: Average Yearly Temperatures in New Haven 1912-1971
>
>
Following are some descriptive statistics of the core R data set nhtemp, a time series of 60 observations recording the mean annual temperature in degrees Fahrenheit in New Haven, Connecticut, between 1912 and 1971.

```{r, echo=FALSE}
hist(nhtemp, 
     main="Frequency of Average Yearly Temperatures", 
     xlab="temperature")
```

You can also embed plots, for example:

```{r, echo=FALSE}
plot(nhtemp)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

To cite out of bibliography:
bibliography:
    - Citations.bib
