---
title: "Assignment 1 Final Output"
output: pdf_document
---

This a summary of the descriptive statistics from Diego and Isabelle's first assignment for the Collaborative Social Science Data Analysis class.

## *Table of Contents*
#[Data Set 1: Average Yearly Temperatures in New Haven 1912-1971](#dataset1)
#[Data Set 2: Violent Crime Rates by US State](#dataset2)

\pagebreak

#Data Set 1: Average Yearly Temperatures in New Haven 1912-1971
>
>
<<<<<<< HEAD
###Introduction to the DataSet
The core R data set nhtemp, titled "Average Yearly Temperatures in New Haven" is a time series of 60 observations recording the mean annual temperature in degrees Fahrenheit in New Haven, Connecticut, between 1912 and 1971.

###Descriptive Statistics

If we look at the summary statistics, we learn that the average temperature has fallen within a small range for the duration of the 60 years. Otherwise, it tells us very little about the actual climate or any change in climate over time, as the temperatures are aggregated by year and not by month or season.
```{r}
summary(nhtemp)
```

However, if we look at the data over time, we can see a general upwards trend in the average temperatures, although they vary considerably from one year to the next.
```{r, echo=FALSE}
plot(nhtemp,
    main="Average Yearly Temperature 1912-1971",
    ylab="Temperature (Fahrenheit)")
```


=======
Following are some descriptive statistics of the core R data set nhtemp, a time series of 60 observations recording the mean annual temperature in degrees Fahrenheit in New Haven, Connecticut, between 1912 and 1971.

```{r, echo=FALSE}
hist(nhtemp, 
     main="Frequency of Average Yearly Temperatures", 
     xlab="temperature")
```

You can also embed plots, for example:

```{r, echo=FALSE}
plot(nhtemp)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
>>>>>>> master

To cite out of bibliography:
bibliography:
    - Citations.bib
