---
title: "Assignment 1 Final Output"
output: 
    pdf_document:
        toc: true
---

This a summary of the descriptive statistics from Diego and Isabelle's first assignment for the Collaborative Social Science Data Analysis class.

\pagebreak

# Data Set 1: Average Yearly Temperatures in New Haven 1912-1971

### Introduction to the DataSet
The core R data set nhtemp, titled "Average Yearly Temperatures in New Haven" is a time series of 60 observations recording the mean annual temperature in degrees Fahrenheit in New Haven, Connecticut, between 1912 and 1971.

### Descriptive Statistics

If we look at the summary statistics, we learn that the average temperature has fallen within a small range for the duration of the 60 years. Otherwise, it tells us very little about the actual climate or any change in climate over time, as the temperatures are aggregated by year and not by month or season.

```{r}
summary(nhtemp)
```

However, if we look at the data over time, we can see a general upwards trend in the average temperatures, although they vary considerably from one year to the next.

```{r, echo=FALSE}
plot(nhtemp,
    main="Average Yearly Temperature 1912-1971",
    ylab="Temperature (Fahrenheit)")
```


\pagebreak

The histogram shows us that most years cluster towards an average temperature of 50-52.

```{r, echo=FALSE}
hist(nhtemp, 
     main="Frequency of Average Yearly Temperatures", 
     xlab="Temperature")
```

\pagebreak

# Data Set 2: Violent Crime Rates by US State

### Introduction to the DataSet

The dataset USArrests contains statistics, in arrests per 100,000 residents for assault, murder, and rape in each of the 50 US states in 1973. It also provides the percent of the population living in urban areas.

There are 50 observations on the 4 following variables:

-Murder (murder arrests per 100,000)

-Assault (assault arrests per 100,000)

-UrbanPop (percent urban population per state)

-Rape (rape arrests per 100,000)

###Descriptive Statistics

Firstly, we use the summary function to have a look at the values of the mean and the median (measures of central tendency) for each of the 4 variables. For murder arrests per 100,000 population, the mean is 7.788 and the median is 7.250.  

```{r}
summary(USArrests$Murder)
```

For assault arrests per 100,000 population, the mean is 170.8 and the median is 159. 

```{r}
summary(USArrests$Assault)
```

For percent of urban population in each state, the mean is 65.54% and the median is 66%. 

```{r}
summary(USArrests$UrbanPop)
```

For rape arrests per 100,000 population, the mean is 21.23 and the median is 20.10. 

```{r}
summary(USArrests$Rape)
```

\pagebreak

Secondly, we generate histograms in order to have a general idea about the distribution of the 4 variables contained in our dataset. For murder arrests per 100,000 population, the histogram tells us that the largest group of US states (12 states) has between 2 and 5 arrests for murder per 100,000 residents. There??s only 1 state that has less than 2 arrests for murder per 100,000 residents and more than 16 states have between 6 and 10 arrests per 100,000 residents (two bars of the histogram).


```{r, echo=FALSE}
hist(USArrests$Murder,
     main = "Arrests for Murder across US states",
     ylab = "Number of US states",
     xlab = "Arrests for Murder per 100k residents")
```

\pagebreak

For assault arrests per 100,000 residents, the histogram tells us that the largest group of states (12 states) has between 100 and 150 arrests for assault per 100,000 residents. The second largest group of states (10 states) has between 250 and 300 arrests for assault per 100,000 residents. On the other hand, while 3 states only have between 0 and 50 arrests for assault per 100,000 residents, 2 states have between 300 and 350. 

```{r, echo=FALSE}
hist(USArrests$Assault,
     main = "Arrests for Assault across US states",
     ylab = "Number of US states",
     xlab = "Arrests for Assault per 100k residents")
```

\pagebreak

```{r, echo=FALSE}
hist(USArrests$UrbanPop,
     main = "Share of Urban Population across US states",
     ylab = "Number of US states",
     xlab = "Share of Urban Population (in %)")
```

```{r, echo=FALSE}
hist(USArrests$Rape,
     main = "Arrests for Rape across US states",
     ylab = "Number of US states",
     xlab = "Arrests for Rape per 100k residents")
```

\pagebreak

```{r, echo=FALSE}
boxplot(USArrests$Murder,
        main = "Murder Arrests",
        ylab = "Arrests per 100k Inhabitants")
```

```{r, echo=FALSE}
boxplot(USArrests$Assault,
        main = "Assault Arrests",
        ylab = "Arrests per 100k Inhabitants")
```

```{r, echo=FALSE}
boxplot(USArrests$Rape,
        main = "Rape Arrests",
        ylab = "Arrests per 100k Inhabitants")
```

\pagebreak

```{r, echo=FALSE}
library(ggplot2)
ggplot2::ggplot(USArrests, aes(Murder, Assault)) + geom_point() + geom_smooth() + theme_bw() + ggtitle("Murder arrests-Assault arrests Correlation") +
  xlab("Murder arrests per 100k habitants") +
  ylab("Assault arrests per 100k habitants")
```

```{r}
cor.test(USArrests$Murder, USArrests$Assault)
```

\pagebreak

```{r, echo=FALSE}
library(ggplot2)
ggplot2::ggplot(USArrests, aes(Murder, Rape)) + geom_point() + geom_smooth() + theme_bw() + ggtitle("Murder arrests-Rape arrests Correlation") +
  xlab("Murder arrests per 100k habitants") +
  ylab("Rape arrests per 100k habitants")
```

```{r}
cor.test(USArrests$Murder, USArrests$Rape)
```

\pagebreak

```{r, echo=FALSE}
library(ggplot2)
ggplot2::ggplot(USArrests, aes(Assault, Rape)) + geom_point() + geom_smooth() + theme_bw() + ggtitle("Assault arrests-Rape arrests Correlation") +
  xlab("Assault arrests per 100k habitants") +
  ylab("Rape arrests per 100k habitants")
```

```{r}
cor.test(USArrests$Assault, USArrests$Rape)
```
